---
title: "p8105_hw2_fc2691"
author: "FC"
date: "10/3/2021"
output: github_document
---

## Problem 1  

This problem uses the Mr. Trash Wheel dataset, available as an Excel file on the course website.  

Read and clean the Mr. Trash Wheel sheet:

- specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel
- use reasonable variable names
- omit rows that do not include dumpster-specific data
- round the number of sports balls to the nearest integer  

Read and clean precipitation data for 2018 and 2019. For each, omit rows without precipitation data and add a variable for year. Next, combine precipitation datasets and convert month to a character variable (the variable month.name is built into R and should be useful).

Write a paragraph about these data; you are encouraged to use inline R. Be sure to note the number of observations in both resulting datasets, and give examples of key variables. For available data, what was the total precipitation in 2018? What was the median number of sports balls in a dumpster in 2019?

```{r}
library(readxl)
library(tidyverse)
library(dplyr)
library(tidyr)

#Clean the data in Mr. Trash Wheel Sheet. 
mr_trash_wheel_sheet =
  readxl::read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", range = "A2:N534") %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls,0))

mr_trash_wheel_sheet

#clean precipitation data for 2018 and 2019

#Reorganize precipitation data for 2018
precipitation_data_2018 = 
  readxl::read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "2018 Precipitation", range = "A2:B15") %>% 
  janitor::clean_names() %>%
  drop_na() %>% 
  mutate(year = 2018)

##Reorganize precipitation data for 2019
precipitation_data_2019 = 
  readxl::read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "2019 Precipitation", range = "A2:B15") %>% 
  janitor::clean_names() %>%
  drop_na() %>% 
  mutate(year = 2019)

precipitation_data_clean = 
  full_join(
    precipitation_data_2018, precipitation_data_2019
  ) %>% 
  mutate(month = month.name[month]) %>% 
  relocate(year, month, total)

precipitation_data_clean

```

In Mr.Trash Wheel sheet, there are data from 2014 to 2021 with 14 variables and 453 observations. In the joint precipitation data for 2018 and 2019, there are 24 observations and 3 variables  Among those data, it could be concluded that the total precipitation in 2018 is `r with(precipitation_data_clean, sum(total[year == 2018]))`. The median number of sports balls in dumpster in 2019 is `r with(mr_trash_wheel_sheet, median(sports_balls[year == 2019]))`.

## Problem 2  

This problem uses the *FiveThirtyEight* data; these data were gathered to create the interactive graphic on this page.  

- First, clean the data in pols-month.csv. Use separate() to break up the variable mon into integer variables year, month, and day; replace month number with month name; create a president variable taking values gop and dem, and remove prez_dem and prez_gop; and remove the day variable.  
- Second, clean the data in snp.csv using a similar process to the above. For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.
- Third, tidy the unemployment data so that it can be merged with the previous datasets. This process will involve switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key variables take the same values.
- Join the datasets by merging snp into pols, and merging unemployment into the result.



```{r echo = FALSE}
#Clean the data in pols-month.csv

library(tidyr)
pols_month_number = 
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(
    col = mon, into = c("year", "month", "day"), sep = "-"
    ) %>% 
  mutate(month = month.abb[as.numeric(month)]) %>%
  mutate(president = ifelse(prez_gop == 1, "gop", "dem")) %>% 
  select(-prez_gop, -prez_dem, -day)
pols_month_number

#Clean the data in snp.csv

snp_stock =
  read_csv("./data/fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>% 
  # mutate(date = as.Date(date, "%m/%d/%y")) %>%
  separate(
    col = date, into = c("month", "day", "year"), sep = "/"
    ) %>% 
  mutate(month = month.abb[as.numeric(month)], 
         close = round(close, 1)) %>% 
  mutate(
    year = as.character(
      ifelse(as.numeric(year) > 21, as.numeric(year) + 1900, as.numeric(year) + 2000))
    ) %>% 
  arrange(year, month) %>% 
  relocate(year, month, everything()) %>% 
  select(-day)
snp_stock

# Tidy the unemployment data so that it can be merged with the previous datasets.

unemployment_percentage =
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>%
  pivot_longer(
    Jan:Dec, 
    names_to = "month", 
    values_to = "unemployment_data"
  ) %>% 
  janitor::clean_names() %>%  
  mutate(year = as.character(year))
unemployment_percentage 

#Join the datasets by merging snp into pols, and merging unemployment into the result.

pols_snp_unemployment_merge = left_join(
  left_join(pols_month_number, snp_stock, by = c("year", "month")), 
  unemployment_percentage, by = c("year", "month"))

pols_snp_unemployment_merge
```

pols-month.csv decribes the number of national politicians who are democratic or repunlican at any given time which including 822 observations and 9 variables. snp.csv describes the closing values of the Standard & Poor’s stock market index on the associated date with 787 observations and 3 variables. During tidy process, the date was seperate to specific year and month. Unemployment.csv describes the percentage of unemployment in each month of the associated year. While changing months from columns to rows, it includes 816 observations. After joining these three tables, the final dimension of table is (`r dim(pols_snp_unemployment_merge)`), the range of years is [`r range(pull(pols_snp_unemployment_merge, year))`]. The tables were joined using key variables year and month. 

## Problem 3  

Load and tidy the data.   
Produce a well-structured, reader-friendly table showing the rank in popularity of the name “Olivia” as a female baby name over time; this should have rows for ethnicities and columns for year. Produce a similar table showing the most popular name among male children over time.

Finally, for male, white non-hispanic children born in 2016, produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis).
```{r}

# Load the data and remove the duplicates. 
library(ggplot2)
popular_baby_names_base =
  read_csv("./data/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>% 
  mutate(
    gender = str_to_title(gender), 
    ethnicity = str_to_title(ethnicity), 
    childs_first_name = str_to_title(childs_first_name)
    ) %>% 
  distinct()

popular_olivia_names = popular_baby_names_base %>% 
  #Showing the rank in popularity of the name “Olivia” as a female baby name over time
  filter(gender == "Female" & childs_first_name == "Olivia") %>% 
  select(-gender, -childs_first_name) %>%
  pivot_wider(
    names_from = c(ethnicity),
    values_from = c(year_of_birth)
    ) %>% 
  relocate(rank, count, everything()) %>% 
  knitr::kable(caption = "The popularity of the name 'Olivia' as a female baby over time")
popular_olivia_names

popular_male_baby_names_count = popular_baby_names_base %>% 
  filter(gender == "Male") %>% 
  group_by(childs_first_name) %>%
  summarise(
    total_count = sum(count), 
    rank_avg = mean(rank)
    ) %>% 
  arrange(rank_avg, desc(total_count))
popular_male_baby_names_count

# From the table above, "Ethan" seems like the most popular name among male children over time. 
popularity_ethan_male_names = popular_baby_names_base %>% 
  filter(gender == "Male" & childs_first_name == "Ethan") %>% 
  select(-gender, -childs_first_name) %>%
  pivot_wider(
    names_from = c(ethnicity),
    values_from = c(year_of_birth)
    ) %>% 
  relocate(rank, count, everything()) %>% 
  knitr::kable(caption = "The popularity of the name 'Ethan' as a male baby over time")
popularity_ethan_male_names

popular_baby_names_base

scatter_plot_male_white_non_hispanic = popular_baby_names_base %>% 
  filter(gender == "Male" & ethnicity == "White Non Hispanic" & year_of_birth == 2016) %>% 
  select(-year_of_birth, -gender, -ethnicity)
scatter_plot_male_white_non_hispanic 
ggplot(scatter_plot_male_white_non_hispanic, aes(x = rank, y = count)) + geom_point()
``` 



